---
title: "Cricket data analysis"
author: "Damodar"
date: "2 December 2018"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(knitr)
library(tidyverse)

```

```{r}
odi = read.csv("C:/Users/Administrator/Desktop/datasets-master/odi-batting.csv")
```



## Data Preview
```{r}
kable(head(odi))
```

### Display the dimension of the data (no. of rows and columns)
```{r}
dim(odi)
```

### Display unique countries and also the total number of unique countries
```{r}
print (unique(odi$Country))
print (nlevels(odi$Country))
```


### Identify how many records are there for Sachin R Tendulkar
```{r}
odi %>% filter(Player == "Sachin R Tendulkar") %>% nrow() #summarise(matches=n())
```

### Display top ten players based on total runs
```{r}
odi %>% group_by(Player) %>% summarise(total_runs=sum(Runs, na.rm=T)) %>% arrange(-total_runs) %>% head(10)
```

##### Identify top ten players by total number of matches and filter the data for those ten players and count the number of rows
```{r}
# Identify top ten players by total numher of matches
top_players = odi %>% group_by(Player) %>% summarise(matches=n()) %>% arrange(-matches) %>% head(10)
top_players

# Filter the data for those ten players and calculate total number of rows
odi %>% filter(Player %in% top_players$Player) %>% nrow()
```

###### For top ten players based on maximum runs and calculate the following 
- total runs
- average score rate
- total number of matches they have played
- total number of ducks

Sort the output by total number of matches
```{r}
top_player = odi %>% group_by(Player) %>% summarise(maxruns=max(Runs, na.rm=T)) %>% arrange(-maxruns) %>% head(10)

odi %>% filter(Player %in% top_player$Player) %>% group_by(Player) %>% summarise(total=sum(Runs, na.rm=T), avgSR=mean(ScoreRate, na.rm=T), total_matches=n(), ducks=sum(Runs==0,na.rm=T), centuries=sum(Runs>99, na.rm=T)) %>% arrange(-total_matches)
```


### Missing values
```{r} 
missing_rows_runs = row.names(odi[is.na(odi$Runs), ])
View(odi[missing_rows_runs, c('Runs', 'ScoreRate')])

View(odi[c(1033), ])
```


### Get percentage of missing values
```{r}
sum(is.na(odi$Runs)) / nrow(odi) * 100
```


### Percentage of outliers
```{r}
q1 = quantile(odi$Runs, 0.25, na.rm=T)
q2 = quantile(odi$Runs, 0.5, na.rm=T)
q3 = quantile(odi$Runs, 0.75, na.rm=T)
val_2per = quantile(odi$Runs, 0.02, na.rm=T) 
val_98per = quantile(odi$Runs, 0.98, na.rm=T)
print(c(q1, q2, q3))
iqr = q3 - q1
iqr
lw = q1 - 1.5 * iqr
uw = q3 + 1.5 * iqr
print (c(lw, uw, val_2per, val_98per))
boxplot(odi$Runs)
```


```{r}
outliers = odi %>% filter((Runs<lw) | (Runs>uw))
nrow(outliers) / sum(!is.na(odi$Runs)) * 100
```


### Frequency distribution on categorical column
```{r}
odi %>% group_by(Player) %>% summarise(total_matches=n()) %>% arrange(-total_matches)
```

```{r}
ggplot(odi, aes(x=Country)) + geom_bar(fill='steelblue') + coord_flip() 
```




```{r}
odi %>% group_by(Player) %>% summarise(total_matches=n(), total_runs=sum(Runs, na.rm=T)) %>% arrange(-total_matches) %>% head(10) %>% ggplot(aes(x=reorder(Player, total_matches), y=total_matches)) + geom_bar(stat='identity', fill='steelblue') + coord_flip() + xlab('Player Name') + ylab('Total No. of Matches')
```



```{r}
odi %>% group_by(Player) %>% summarise(total_matches=n() / nrow(odi) * 100) %>% arrange(-total_matches) %>% mutate(cum_sum= cumsum(total_matches)) %>% ggplot(aes(x=reorder(Player, cum_sum), y=cum_sum)) + geom_line(group=1)
```


```{r}

odi$Bins = cut(odi$Runs, breaks =c(-0.01,10, 49, 75, 90, 99, Inf), labels=c('b1', 'b2', 'b3','b4','b5', 'b6') )
odi %>% ggplot(aes(x=Bins)) + geom_bar(fill='steelblue')
```




### Date columns
```{r}
odi$date = as.Date(odi$MatchDate, '%m-%d-%Y')
odi$year = format(odi$date, '%Y')
odi$month = format(odi$date, '%m')
odi$day = format(odi$date, '%d')
odi$weekday = format(odi$date, '%A')
odi$quarter = quarters(odi$date)
odi$quarter = format(odi$date, '%q') # Doesn't work. Need to check
View(odi %>% select(date, year, month, day, weekday, quarter))
```

## Bi-Variate Analysis
```{r}
odi %>% filter(Player=='Sachin R Tendulkar') %>% group_by(year) %>% summarise(total=sum(Runs, na.rm=T), avg_runs=mean(Runs, na.rm=T), median_runs=median(Runs, na.rm=T), matches=n()) %>% ggplot(aes(x=year, y=matches)) + geom_line(group=1, color='steelblue')
```

### Bivariate Analysis
```{r}
odi %>% filter(Player == "Sachin R Tendulkar") %>% ggplot(aes(x=Runs, y=ScoreRate))+geom_point() + geom_smooth()
```


```{r}
odi %>% group_by(Country) %>% summarise(Runs=sum(Runs, na.rm=T), ScoreRate=mean(ScoreRate, na.rm=T)) %>% ggplot(aes(x=Runs, y=ScoreRate)) + geom_point()
```


```{r}
top_players = odi %>% group_by(Player) %>% summarise(Runs=sum(Runs, na.rm=T)) %>% arrange(-Runs) %>% head(5)

odi %>% filter(Player %in% top_players$Player) %>% ggplot(aes(x=Player, y=Runs)) + geom_boxplot()
```

```{r}
odi %>% ggplot(aes(x=Country, y=Runs)) +geom_boxplot()
```

